Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression

### 2016-07-29 manuscript.tex
moved phylogeny section

\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1). ). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independence are likely violated, since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate. To address this issue, we used a maximum likelihood approach using the the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio (\X Supp. Figures). Here, \X{...explain more}. Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males. As no significant difference was observed (\X Supp. fig.), we report only results from females.

@@

We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}. To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data. To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$. We then calculated the average \theta over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure). Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).

@@

>>>>

\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated in our , since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate. To address this issue, we used a maximum likelihood approach using the the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio (\X Supp. Figures). Here, \X{...explain more}.  
To test for population substructure between the two sex chromosome races, which could affect estimates diversity, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study. The analysis was conducted on an alignment of X- and Y-linked genes from \texit{R. hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \texit{R. bucephalophorus} used to root the tree. Evolutionary relationships of sex chromosome races in Rumex hastatulus, inferred using the Neighbor-Joining method (Saitou and Nei 1987). The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al. 2004). 
Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males. As no significant difference was observed (\X Supp. fig.), we report only results from females.

@@

We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}. To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data. To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$. We then calculated the average $\theta$ over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure). Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).

@@

%In contrast to the Texas sex chromosome race, our analysis revealed that Y-chromosome diversity was considerably higher in the North Carolina race (see Supplementary Material; Figure S1). This difference could have arisen due to population substructure, resulting in an elevated overall Y diversity in the pooled sample. To test this hypothesis, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study. The resulting sequence phylogeny (Figure 4) revealed strong support for the Y-linked genes from the XY1Y2 race being paraphyletic (98% bootstrap support). In particular, our analysis revealed that samples from two populations, both from South Carolina (hereafter the SC sub-clade), formed a monophyletic group, whereas samples from Florida and Georgia (hereafter the FL sub-clade) were also monophyletic, but were more closely related to the XY Texas race. Although our sampling here is limited, these results suggest that introgression may have occurred between the ancestral Texas (XY race) and the derived North Carolina (XY1Y2) race, leading to a second, derived, XY1Y2 sub-clade and an elevated level of Y-diversity in our pooled sample. As these races are inter-fertile (Smith reference about hybrid zone), we suggest that the SC sub-clade likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY race. Notably, when we separated samples by the two inferred sub-clades of Y–linked sequences, the estimated level of Y diversity was consistent across all three clades, suggesting that the extensive reduction in diversity on this chromosome is not a population or race-specific phenomenon (see Supplementary Material, Figure S1).

%figure

%Figure 4. Evolutionary relationships of sex chromosome races in Rumex hastatulus, inferred using the Neighbor-Joining method (Saitou and Nei 1987). The percentage of replicate trees in which the associated sequences clustered together in the bootstrap test (1000 replicates) is shown next to branches (Felsenstein 1985). The phylogeny is drawn to scale, with branch lengths in the same units as those of the evolutionary distances used to infer the tree. The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al. 2004) and are in the units of the number of base substitutions per site. The analysis was conducted on an alignment of X- and Y-linked genes from Rumex hastatulus, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species Rumex bucephalophorus used to root the tree. ‘N’ designates the North Carolina race, and ‘T’ the Texas race, with numbers 1-6 corresponding to each of the 6 populations sampled. The inferred SC (South Carolina), and FL (Florida) sub-clades are indicated.
